---
title: "Coding Challenge"
author: Sahil Hooda
output:
  word_document: default
  html_document: default
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```

#Requiring some important libraries
```{r}
library(ggplot2)
library(dplyr)
library(readr)
library(mosaic)
library(tidyr)
library(chron)
library(reshape2)
library(ggpubr)
library(caret)
library(purrr)
library(pROC)
```

##Loading the Data into R
```{r}
caredashmc <- read.csv("/Users/sahil/Desktop/CareDash/Data.csv")
```

```{r}
sum(as.logical(caredashmc$sends))
sum(as.logical(caredashmc$opens))
sum(as.logical(caredashmc$clicks))
sum(as.logical(caredashmc$unsubs))
```
# There are 296,174 sends, 74,923 opens, 5139 clicks, 115 unsubs

#Removing unsubs from clicks 
```{r}
caredashmc$opens <- caredashmc$opens - caredashmc$unsubs
caredashmc$clicks <- caredashmc$clicks - caredashmc$unsubs
```

#Separating Date & Time
```{r}
caredashmc <- separate(data=caredashmc, col= send_time, c("send_date", "send_time"), sep=" ")
caredashmc <- separate(data=caredashmc, col= open_time, c("open_date", "open_time"), sep=" ")
```
#Cleaning up the time Variable
```{r}
#Cleaning time variable up
caredashmc$send_time <- substr(caredashmc$send_time, 1, 8)
caredashmc$open_time <- substr(caredashmc$open_time, 1, 8)
```

#Character to rdate
```{r}
caredashmc$send_date <- as.Date(caredashmc$send_date, format = "%Y-%m-%d")
caredashmc$open_date <- as.Date(caredashmc$open_date, format = "%Y-%m-%d")
caredashmc$send_time <- chron(times. = caredashmc$send_time)
caredashmc$open_time <- chron(times. = caredashmc$open_time)
```

# Extracting hour from send time for grouping
```{r}
caredashmc$send_time_hour <- hours(caredashmc$send_time)
caredashmc$open_time_hour <- hours(caredashmc$open_time)
```

#Tried to fix what the time looks like 
```{r}
caredashmc$send_time_hourpm <- caredashmc$send_time_hour -12
```

```{r}
caredashmc$daydiff <- caredashmc$open_date - caredashmc$send_date
table(caredashmc$daydiff)
```
##The maximum difference between open date and send date is only 8 days. Hence, open date is not a factor to consider in my analysis. 


###SUMMARY STATISTICS
#Finding the range of dates of the campaign
```{r}
range(caredashmc$send_date)
range(caredashmc$open_date, na.rm = TRUE)
```
#The range for dates when E-mails were sent out are 2014-03-17 and 2014-03-23
#The range for dates when E-mails were opened is 2014-03-17 and 2014-03-25
#This helps because now we don't have to worry about fluctuations through month and year

#What day of the week was the email sent and opened? 
```{r}
caredashmc$sday <- weekdays(as.POSIXct(caredashmc$send_date), abbreviate = F)
caredashmc$oday <- weekdays(as.POSIXct(caredashmc$open_date), abbreviate = F)
```

```{r}
caredashmc$sdayntime <- paste(caredashmc$sday, caredashmc$send_time_hourpm)
```

#Number of E-mails sent out by each day
```{r}
dat <- select(caredashmc, 5:8)
daysummary <- aggregate(dat, by=list(Send_day=caredashmc$sday), FUN = sum)
hoursummary <- aggregate(dat, by=list(Send_hour=caredashmc$send_time_hourpm), FUN = sum)
daynhoursummary <- aggregate(dat, by=list(Send_daynhour=caredashmc$sdayntime), FUN = sum)
```

## VISUALIZING THE DATA

#sendsperday
```{r}
#plot1 <- ggplot(data = caredashmc, aes(x = sday, y=sends)) +
#  geom_histogram(stat = 'identity') +
#  theme(text = element_text(size=20),
#        axis.text.x = element_text(angle=90, hjust=1))
#plot1
```
#Opensperday
```{r}
#plot2 <- ggplot(data = caredashmc, aes(x = sday, y=opens)) +
#  geom_histogram(stat = 'identity') +
#  theme(text = element_text(size=20),
#        axis.text.x = element_text(angle=90, hjust=1))
#plot2
```
#Sends distribution per hour (Graph is bad -> Can't see all data)
```{r}
#plot3 <- ggplot(data = caredashmc, aes(x = send_time_hourpm, y=sends)) +
#  geom_histogram(stat = 'identity') +
#  theme(text = element_text(size=20),
#        axis.text.x = element_text(angle=90, hjust=1))
#plot3
```
#opens distribution per hour (Graph misrepresents data -> due to scale, can't see all data)
```{r}
#plot4 <- ggplot(data = caredashmc, aes(x = send_time_hourpm, y=opens)) +
#  geom_histogram(stat = 'identity') +
#  theme(text = element_text(size=20),
#        axis.text.x = element_text(angle=90, hjust=1))
#plot4 
```

```{r}
plot(hoursummary$Send_hour, hoursummary$sends)
```

## Plot for percentage of opens for sends and percentage of clicks for sends per hour
```{r}
hoursummary <- subset(hoursummary, !(sends<10))

hoursummary$openpercent <- (hoursummary$opens/hoursummary$sends)*100
hoursummary$clickpercent <- (hoursummary$clicks/hoursummary$sends)*100
#hourlyjoint$clickbyopenpercent <- hourlyjoint$Clicks/joint$Opens

newjoint1 <-select(hoursummary, Send_hour, openpercent, clickpercent)

joint.long1<-melt(newjoint1,id.vars="Send_hour")

ggplot(joint.long1,aes(x=Send_hour,y=value,fill=factor(variable)))+
  geom_bar(stat="identity",position="dodge")+
  xlab("Time in PM")+ylab("Percent")
```
```{r}
ggplot(hoursummary,aes(x=Send_hour,y=clickpercent))+
  geom_bar(stat="identity",position="dodge")+
  xlab("Time in PM")+ylab("Click Percent")
```


## Plot for percentage of opens for sends and percentage of clicks for sends per day
```{r}
daysummary$openpercent <- (daysummary$opens/daysummary$sends)*100
daysummary$clickpercent <- (daysummary$clicks/daysummary$sends)*100
#daysummary$clickbyopenpercent <- joint$Clicks/joint$Opens

newjoint <-select(daysummary, Send_day, openpercent, clickpercent)

joint.long<-melt(newjoint,id.vars="Send_day")
ggplot(joint.long,aes(x=Send_day,y=value,fill=factor(variable), legend = factor(variable)))+
  geom_bar(stat="identity",position="dodge")+
  xlab("Day")+ylab("t")
```

```{r}
daynhoursummary$Send_daynhour <- as.factor(daynhoursummary$Send_daynhour)
plot(daynhoursummary$Send_daynhour, log(daynhoursummary$sends))
```
##Plot for percentage of opens and clicks for sends for sends at a day and time
## Removed day and time combinations that had less than 5 sends 
```{r}
daynhoursummary <- subset(daynhoursummary, !(sends<5))

daynhoursummary$openpercent <- (daynhoursummary$opens/daynhoursummary$sends)*100
daynhoursummary$clickpercent <- (daynhoursummary$clicks/daynhoursummary$sends)*100
#daynhrjoint$clickbyopenpercent <- daynhrjoint$Clicks/daynhrjoint$Opens

newjoint2 <-select(daynhoursummary, Send_daynhour, openpercent, clickpercent)

joint.long2<-melt(newjoint2,id.vars="Send_daynhour")
ggplot(joint.long2,aes(x=Send_daynhour,y=value,fill=factor(variable), legend = factor(variable)))+
  geom_bar(stat="identity",position="dodge")+
  theme(text = element_text(size=10),
        axis.text.x = element_text(angle=90, hjust=1)) +
  xlab("Day And Time in PM")+ylab("Percent")
```



```{r}
caredashmc$opens<- as.factor(caredashmc$opens)
logit <- glm(opens ~ sday + send_time_hour + sday:send_time_hour, data = caredashmc, family = "binomial")
anova(logit, test = "Chisq")
```
#The Interaction term helps the model


```{r}
#predict(logit, type = "response")
#logit.pred <- ifelse(logit$fitted.values > 0.5, 1, 0)

#logit_table <- table(logit.pred, caredashmc$opens)
#roc.info = roc(caredashmc$opens, logit$fitted.values, plot = TRUE, legacy.axes = TRUE, 
#                main = "Figure 4", xlab = "False Positive Percentage", ylab = "True Positive Percentage", 
#                print.auc = TRUE)
#logit_table
```

```{r}
pred.logit<- predict(logit, type = "response")
agpred.logit <- aggregate(pred.logit, by = list(caredashmc$sday, caredashmc$send_time_hour), FUN = mean)

agpred.logit <- agpred.logit[(agpred.logit$Group.2>14 & agpred.logit$Group.2<21),]
agpred.logit <- agpred.logit[!(agpred.logit$Group.1=="Sunday" & agpred.logit$Group.2==15),]
agpred.logit <- agpred.logit[!(agpred.logit$Group.1=="Saturday" & agpred.logit$Group.2==15),]

```

```{r}
agpred.logit$Group.1 <- paste(agpred.logit$Group.1, agpred.logit$Group.2)
```

```{r}
ggplot(agpred.logit,aes(x=Group.1,y=x*100),)+
  geom_bar(stat="identity",position="dodge")+
  theme(text = element_text(size=10),
        axis.text.x = element_text(angle=90, hjust=1)) +
  xlab("Day")+ylab("Percent")
```

